Cascades, Product Conversion, Whatever

### 2012-12-12 paper.tex
Added cloud graph, and minor changes

\title{Cascades, Product Conversion, Whatever}
\author{T. Martian \& E. Brinkbot}

@@

Many models with simple spreading rules have been proposed
\cite{Arthur89, Morris00, Watts02} to explain, for example, how
breaking news spreads over the internet or how a new technology (such
as the iPod) spreads in popularity. These spreading rules usually take
one of two forms. The rules either govern a stochastic spreading
process, or they assume that each node in the network is a strategic
agent which makes a decision (e.g. to buy an iPod or to buy a Zune)
and derives greater utility if his neighbors make the same
decision. To the best of our knowledge, all current strategic agent
models make a simplifying assumption that the agents are only
myopically strategic \cite{Chierichetti12}: they make decisions at
each round to maximize their utility at that point in the game. But,
each agent's behavior can have profound impact on the future state of
the network and thus the agent's own final utility. Our interest is in
studying network processes under the more realistic assumption that
agents make decisions conscious of their future influence on outcomes,
in order to maximize their expected final utility.

@@

We propose a cascade spreading model that borrows heavily from the
scheduler mediated cascade process introduced in
\cite{Chierichetti12}. In our model every node in an undirected graph
is a risk neutral rational agent. Each node is also assigned a binary
type which affects their utility, pulled from a binomial
distribution. In this paper we call the types Yes ($Y$) and No
($N$). There is an additional scheduler agent, which, at each round of
the game picks one of the nodes in the graph. The chosen node picks a
``choice'' (also Yes or No) and the agent is committed to that
``choice'' for the rest of the game. The game ends when all nodes have
picked a ``choice.''

We keep the same scheduler behavior as in the previous paper. The
scheduler's utility is based off of the number of nodes that choose Yes
at the end of the game. We assume the scheduler can change its plan
after it sees what a node chooses. We also assume it has full
knowledge of the nodes' strategies, as well as knowledge of the full
graph, and any game parameters. The only aspect the scheduler does not
know is the individual node types.

The nodes' utility is the sum of all of their neighbors that match
their choice plus some constant ($\pi$) if their choice matches their
type. Strategic nodes make decisions to maximize their expected
profit. We assume strategic nodes know the schedulers decision making
(or utility function), the graph structure, and the probability that a
node is assigned Yes type ($p$), but not any other nodes type. To keep
the model interesting, we restrict $p < 0.5$. Conversely, myopic nodes
only try to maximize their current utility, ignoring any nodes that
have not made a decision.

We consider games where either all nodes in the graph are strategic or
all are myopic, and want to see what graphs / game parameters contribute to
significant differences between the expected number of Yesses when
nodes are strategic versus myopic.

We formally specify the model as follows
\begin{align*}
  & G(V,E) && \text{social network} \\
  & |V| = n \\
  & c_i \in \{Y,N\} && \text{choice of node $i$} \\
  & t_i \in \{Y, N\} && \text{type of node $i$} \\
  & p(t_i = Y) = p < 0.5 \\
  & u_i(\mathbf c) = \pi \mathbb I[t_i = c_i] + \sum_{j \in N(i)}
  \mathbb I[c_j = c_i] && \text{node utility} \\
  & sched(\mathbf c) && \text{schedule function}
\end{align*}
where $\mathbb I$ is the indicator function, and $N$ is the neighborhood
function. The schedule function returns the next node to pick given
the choices that have been made so far.

The full game can be described by the following algorithm. The
scheduler picks an optimal first node in the graph. That node chooses
either Yes or No based on all of the parameters and its
knowledge. Based on this result the scheduler picks a new node, and so
on until every node has been picked and the game finishes.

@@

Is seems likely that computing the strategic choices and payouts is a
hard problem, but we will show an algorithm that can compute an
optimal online schedule and calculate $\mathbb E[\#Y|stragetic]$ for arbitrary
block models in time polynomial in the number of blocks.

A graph is a $k$-block model if the vertices can be divided into $k$
disjoint subsets such that every node in the same subset has
incoming and outgoing edges to the same set of nodes. Another way to
view this property is by looking at the adjacency matrix of the
graph. A block model graph can be written as a symmetric block matrix
with $k^2$ blocks. For the sake of the algorithm, we will define
\begin{equation*}
  \{s_i\}_{i=1..k}
\end{equation*}
as the number of nodes in each block.

An example of a 2-block model would be the star graph, which can be
expressed in adjacency matrix form as
\begin{equation*}
  \begin{array}{r}
    1 \{ \\
    n-1 \{
  \end{array}
  \left[ \begin{array}{cc}
      0 & 1 \\
      1 & 0
      \end{array} \right]
\end{equation*}
and has $\mathbf s = \{1, n-1\}$.

@@

The two tables recursively reference each other to calculate their
entries, where each entry is the choice, and the expected number of
Yesses at the end of the game in each block conditioned on that
choice. If an entry is already calculated, then the lookup is constant
time, otherwise the lookup may cause a cascade of calculations, which
is bounded by the polynomial time to compute both tables.\footnote{It
  is possible that instead of looking these up on the fly, the table
  could be precomputed to speed up the complexity slightly, but we
  believe that this speedup would be very slight, and would make the
  algorithm much more complicated}

@@

To determine the optimal block choice for the scheduler, the decision
function only needs to know how many nodes from each block have
already gone and what proportion (or number) of the nodes have chosen
Yes. Because nodes within a block are exchangeable, this information
completely characterizes the state of the game at the time of
scheduler choice.

@@

The scheduler needs to check the possible outcome given it picks a Yes
or a No type node from each of the $k$ blocks that have a most one
node possible to pick. It then marginalizes out the type of the node
to calculate the expected number of Yesses for each choice of
block. The optimal choice is simply the one that maximizes the
expected number of Yesses.

A node needs to check its own expected final utility given it chooses
Yes or No, and choose the choice that maximizes it. The node can query
the scheduler function to determine the expected number of Yesses in
each block conditioned on the node choosing Yes or No. The node then
trivially calculates its expected utility for each choice and picks
the maximum.

Completely calculating the results of these functions is trivially
simple, and in doing so completely finds the optimal online schedule
and optimal node choice along that schedule.

@@

See Section \ref{sec:algorithm} for an explanation of the algorithm.

Lookups are constant time, so the total time complexity is 
 bounded by the number of entries in the table times the number of
table lookups it takes to compute one entry in the table.

@@

We test a star graph, a clique, and a special graph we call a cloud
graph. A cloud graph consists of two singular ``outer'' vertices of
degree $a$ and $b$ respectively, one singular ``inner'' vertex of
degree $a+b$, and two clouds of size $a$ and $b$ respectively, with
each vertex of degree two. Each of the ``outer'' vertices is connected
to every node in a distinct cloud. The singular ``inner'' vertex is
connected to every node in both clouds. This graph, under certain
parameters, provably obtains more $Y$ adoptions for strategic users.

@@

Our results are interesting and certainly show that the difference
between optimal play with myopic and strategic users is a complex
space even very simple block graphs. However, there are still a number
of interesting unexplored questions that we have not looked at in the
scope of the problem.

Despite the fact that obvious patterns don't exist in the performance
ratio, it is still possible that the performance ratio can be bounded
in certain circumstance beyond the star graph. Proving bounds on the
performance ratio could help characterize the discontinuous regions in
the plots where myopic superiority switches in favor of strategic
agents.

>>>>

\title{Cascades, Product Conversion, Whatever}
\author{T. Martin \& E. Brinkman}

@@

Many models with simple spreading rules have been proposed
\cite{Arthur89, Morris00, Watts02} to explain, for example, how
breaking news spreads over the internet or how a new technology (such
as the iPod) spreads in popularity. These spreading rules usually take
one of two forms. The rules either govern a stochastic spreading
process, or they assume that each node in the network is a strategic
agent which makes a decision (e.g. to buy an iPod or to buy a Zune)
and derives greater utility if her neighbors make the same
decision. To the best of our knowledge, all current strategic agent
models make a simplifying assumption that the agents are only
myopically strategic \cite{Chierichetti12}: they make decisions at
each round to maximize their utility at that point in the game. But,
each agent's behavior can have profound impact on the future state of
the network and thus the agent's own final utility. Our interest is in
studying network processes under the more realistic assumption that
agents make decisions conscious of their future influence on outcomes,
in order to maximize their expected final utility.

@@

We propose a cascade spreading model that borrows heavily from the
scheduler mediated cascade process introduced in
\cite{Chierichetti12}. In our model every node in an undirected graph
is a risk neutral rational agent. Each node is also assigned a binary
type which affects their utility, sampled from a Bernoulli
distribution. In this paper we call the types Yes ($Y$) and No
($N$). There is an additional scheduler agent, which, at each round of
the game picks one of the nodes in the graph. The chosen node picks a
``choice'' (also Yes or No) and the agent is committed to that
choice for the rest of the game. The game ends when all nodes have
picked a choice.

We keep the same scheduler behavior as in the previous paper. The
scheduler's utility is maximized by maximizing the number of nodes
that choose Yes at the end of the game. We assume the scheduler can
change its plan after it sees what a node chooses. We also assume the
scheduler has full knowledge of the nodes' strategies, as well as
knowledge of the full graph, and any game parameters. The only aspect
the scheduler does not know is the realized individual node types.

A node's utility is the sum of all of their neighbors that match their
choice plus some constant ($\pi$) if their choice matches their
type. Strategic nodes make decisions to maximize their expected profit
at the end of the game. We assume strategic nodes know the scheduler's
decision making (or utility function), the graph structure, and the
probability that a node is assigned Yes type ($p$), but not any other
node's type. To keep the model interesting, we restrict $p <
0.5$. Conversely, myopic nodes only try to maximize their current
utility, ignoring any nodes that have not made a decision.

We consider games where either all nodes in the graph are strategic or
all are myopic, and want to see what graphs or game parameters
contribute to significant differences between the expected number of
Yeses when nodes are strategic versus myopic.

We formally specify the model as follows
\begin{align*}
  & G(V,E) && \text{social network} \\
  & |V| = n \\
  & c_i \in \{Y,N\} && \text{choice of node $i$} \\
  & t_i \in \{Y, N\} && \text{type of node $i$} \\
  & P(t_i = Y) = p < 0.5 \\
  & u_i(\mathbf c) = \pi \mathbb I[t_i = c_i] + \sum_{j \in N(i)}
  \mathbb I[c_j = c_i] && \text{node utility} \\
  & sched(\mathbf c) && \text{schedule function}
\end{align*}
where $\mathbb I$ is the indicator function, and $N$ is the neighborhood
function. The schedule function returns the next node to pick given
the choices that have been made so far.

The full game can be described by the following algorithm.
\begin{enumerate}
\item \label{start} The scheduler picks an optimal node in the graph
\item That node chooses either Yes or No based on all of the
  parameters and its knowledge
\item Repeat from step \ref{start} until all nodes have picked a
  choice
\end{enumerate}

@@

Is seems likely that computing the strategic choices and payouts is a
hard problem\footnote{We don't have a proof of this yet, but it is
  interesting future work}, but we will show an algorithm that can
compute an optimal online schedule and calculate $\mathbb
E[\#Y|stragetic]$ for arbitrary blockmodels in time polynomial in the
number of blocks.

A graph is a $k$-blockmodel if the vertices can be divided into $k$
disjoint subsets such that every node in the same subset has
incoming and outgoing edges to the same set of nodes. Another way to
view this property is by looking at the adjacency matrix of the
graph. A block model graph can be written as a symmetric block matrix
with $k^2$ blocks. For the sake of the algorithm, we will define
\begin{equation*}
  \{s_i\}_{i=1..k}
\end{equation*}
as the number of nodes in each block.

An example of a 2-blockmodel would be the star graph, which can be
expressed in adjacency matrix form as
\begin{equation*}
  \begin{array}{r}
    1 \{ \\
    n-1 \{
  \end{array}
  \left[ \begin{array}{cc}
      0 & 1 \\
      1 & 0
      \end{array} \right]
\end{equation*}
and has $\mathbf s = \{1, n-1\}$.

@@

The two tables recursively reference each other to calculate their
entries, where each entry is the choice, and the expected number of
Yeses at the end of the game in each block conditioned on that
choice. If an entry is already calculated, then the lookup is constant
time, otherwise the lookup may cause a cascade of calculations, which
is bounded by the polynomial time to compute both tables.\footnote{It
  is possible that instead of looking these up on the fly, the table
  could be precomputed to speed up the complexity slightly, but we
  believe that this speedup would be very slight, and would make the
  algorithm much more complicated}

@@

To determine the optimal block choice for the scheduler, the decision
function only needs to know how many nodes from each block have
already gone and how many of the nodes have chosen Yes. Because nodes
within a block are exchangeable, this information completely
characterizes the state of the game at the time of scheduler choice.

@@

The scheduler needs to check the possible outcome given it picks a Yes
or a No type node from each of the $k$ blocks that have a most one
node possible to pick. It then marginalizes out the type of the node
to calculate the expected number of Yeses for each choice of
block. The optimal choice is simply the one that maximizes the
expected number of Yeses.

A node needs to check its own expected final utility given it chooses
Yes or No, and choose the choice that maximizes it. The node can query
the scheduler function to determine the expected number of Yeses in
each block conditioned on the node choosing Yes or No. The node then
trivially calculates its expected utility for each choice and picks
the maximum.

Completely calculating the results of these functions is simple, and
in doing so computes the optimal online schedule and optimal
node choice along that schedule.

It is trivial to see that this algorithm can also solve the
myopic case, as well as directed weighted arbitrary $k$-blockmodels.

@@

Lookups in the tables are constant time, so the total time complexity is 
 bounded by the number of entries in the table times the number of
table lookups it takes to compute one entry in the table.

@@

We test a star graph, a clique, and a special graph we call a cloud
graph (See Figure~\ref{fig:cloud_graph}). A cloud graph consists of two singular ``outer'' vertices of
degree $a$ and $b$ respectively, one singular ``inner'' vertex of
degree $a+b$, and two clouds of size $a$ and $b$ respectively, with
each vertex of degree two. Each of the ``outer'' vertices is connected
to every node in a distinct cloud. The singular ``inner'' vertex is
connected to every node in both clouds. This graph, under certain
parameters, provably obtains more $Y$ adoptions for strategic users.

\begin{figure}
  \centering
  \includegraphics[width=0.6\textwidth]{cloud_graph.pdf}
  \caption{Stylized Cloud Graph.}
  \label{fig:cloud_graph}
\end{figure}

@@

Our results show that the difference between optimal play with myopic
and strategic users is a complex space even for very simple block
graphs. However, there are still a number of interesting unexplored
questions that we have not looked at in the scope of the problem.

Despite the fact that obvious patterns do not exist in the performance
ratio, it is still possible that the performance ratio can be bounded
in certain circumstance beyond the star graph. Proving bounds on the
performance ratio could help characterize the discontinuous regions in
the plots where myopic superiority switches in favor of strategic
agents.
